---
title: "607 Assignment1: Loading Data into a Data Frame "
author: "Mahmud Hasan Al Raji"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Presidential Primary Polls files contain data since the most recent election.
Link to the data set is <https://projects.fivethirtyeight.com/polls-page/data/president_primary_polls.csv>

## Load libraries
Required libraries

```{r message=FALSE, warning=FALSE}

library(dplyr)
library(magrittr)
library(RCurl)
```

## Import file as CSV

Data has header row so we must indicate “header=T”
```{r import data}


x <- getURL("https://projects.fivethirtyeight.com/polls-page/data/president_primary_polls.csv")
df<-data.frame(read.csv(text=x, header=T))

```

## Selection  of Variables

Check column headers
```{r }

df1=df %>% select(poll_id,pollster_id,pollster,pollster_rating_id,fte_grade,methodology,state,population,subpopulation,sample_size,question_id,cycle,office_type,stage,party,answer,candidate_id,candidate_name,pct)

head(df1)

```


## Rename column headers

Check renamed column headers
```{r }
df1 %<>% rename(fivethirtyeight_grade=fte_grade,percentage=pct)
head(df1)

```


## Transformation of abbreviated values

Check transformed values in columns
```{r }
df1 %<>% mutate(population=case_when(population=="lv"~"likely voters",
                                          population=="rv"~"registered voters",
                                          population=="a"~"all adults",
                                          population=="v"~"voters"),
             subpopulation=case_when(subpopulation=="r"~"republican",
                                     subpopulation=="d"~"democratic" ))
head(df1)
```

## Select a subset by column and row
prepare a table showing population and party wise candidate name who got over 55% vote in presidential primary polls
```{r } 
table1<-subset(df1,percentage>55,select=c(state,population,party,candidate_name,percentage))
table1
```
# Conclusion 
Presidential Primary Polls data is used here to identify population support and party wise candidate names who achieved more than 55% vote in different states in Presidential Primary Polls.
